# Reproducible Research: Peer Assessment 1


## 1. Loading and preprocessing the data

### 1.1. Loading the data
```{r 1_1}
require(data.table)
require(ggplot2)
require(scales)
setwd("~/Git/RepData_PeerAssessment1/")
unzip("./activity.zip", "activity.csv")
dtData <- fread("activity.csv")
```

### 1.2. Process/transform the data
```{r 1_2}
dtData$date <- as.POSIXct(strptime(dtData$date, format = "%Y-%m-%d"))
dtData$weekend <- as.factor(weekdays(dtData$date) %in% c("Sunday","Saturday"))

dtPerDay <- dtData[,list(sum = sum(steps),mean = mean(steps)), by = date]
dtPerDay <- cbind(dtPerDay,tapply(dtData$steps, dtData$date, median))
setnames(dtPerDay, c("V2"), c("median"))
```

## 2. What is mean total number of steps taken per day?

### 2.1. Figure: Total number of steps taken each day
```{r 2_1}
ggplot(dtPerDay, aes(x=dtPerDay$sum)) +
    geom_histogram(binwidth=500) +
    xlab("steps per day") +
    ylab("count of days") +
    ggtitle("Total number of steps taken each day") +
    scale_x_continuous(breaks=seq(0,max(dtPerDay$sum, na.rm=TRUE), by=2000))
```

### 2.2. Table: Mean and median for total number or steps taken per day 
```{r 2_2}
dtPerDay[,list(date=date,mean=mean,meadian=median)]
```


## 3. What is the average daily activity pattern?

### 3.1. Daily pattern
```{r 3_1}
dtPerInterval <- dtData[, list(sum = sum(steps,na.rm=TRUE), mean = mean(steps,na.rm=TRUE)), by = interval]
ggplot(dtPerInterval, aes(x=dtPerInterval$interval, y=dtPerInterval$mean)) +
  geom_line() +
  xlab("Time of day") +
  ylab("Average number of steps") +
  ggtitle("Number of steps taken, averaged across all days")
```

### 3.2. Maximum activity
```{r 3_2}
maxInterval <- dtPerInterval[max(dtPerInterval$mean)==dtPerInterval$mean,interval]
```
Interval: `r maxInterval` contains maximum number of steps on average across all measured days. 

## 4. Imputing missing values

### 4.1. Report missing values
```{r 4_1}
misValCount <- sum(!complete.cases(dtData))
```
There are `r misValCount` records with missing values in the dataset.

### 4.2. Strategy for filling the missing values
We will use mean of the mean across all days for missing values.
```{r 4_2}
dtMerged <- merge(dtData[is.na(dtData$steps),],dtPerInterval, by="interval")
```

### 4.3. Filling the missing values in new data.frame
```{r 4_3}
dtFilled <- dtData
dtFilled[is.na(dtFilled$steps)]$steps <- as.integer(dtMerged$mean)
dtFilledPerDay <- dtFilled[,list(sum = sum(steps),mean = mean(steps)), by = date]
dtFilledPerDay <- cbind(dtFilledPerDay,tapply(dtFilled$steps, dtFilled$date, median))
setnames(dtFilledPerDay, c("date","sum","mean","V2"), c("date","sum","mean","median"))
```

### 4.4. Visualize with filled values
```{r 4_4_1}
ggplot(dtFilledPerDay, aes(x=dtFilledPerDay$sum)) +
    geom_histogram(binwidth=500) +
    xlab("steps per day") +
    ylab("count of days") +
    ggtitle("Total number of steps taken each day") +
    scale_x_continuous(breaks=seq(0,max(dtFilledPerDay$sum, na.rm=TRUE), by=2000))
```

```{r 4_4_2}
dtFilledPerDay[,list(date=date,mean=mean,meadian=median)]
```
**Observation:**  
- Filling means across all days for the missing values slightly changes the histogram, we see there are overall more days shown in the plot.
- The difference is also visible in the table of *MEANS* and *MEDIANS*, while there were missing values in the secion No. 2, we can see values with artificialy inserted data. Especially notable is diffrence in *MEDIANS*.

## 5.Are there differences in activity patterns between weekdays and weekends? 

### 5.1 Set weekend factor levels
```{r 5_1}
levels(dtFilled$weekend) <- c("WEEKDAY", "WEEKEND")
```

### 5.2 Visualize weekends vs weekdays
```{r 5_2}
ggplot(dtFilled, aes(x=dtFilled$interval, y=dtFilled$steps)) +
  stat_summary(fun.y=mean, geom="line") +
  xlab("Time of day") +
  ylab("Average number of steps") +
  ggtitle("Number of steps taken, averaged across all days") + facet_grid(weekend ~ .)
```